Vagueness as Cost Reduction

### 2016-08-03 papers/2016/vcr.tex
no longer tracking contribution

\documentclass[
a4paper 
%, man
, doc
%, jou
, draftfirst 
, longtable
]{apa6} 
\usepackage{rotating} 
                          % use this in jou and doc modes only, with \sideways. 
                          % In man mode use the documentclass option longtable and do not load rotating.
\usepackage{csquotes}  
\usepackage[american]{babel}                     
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, doi=false, isbn=false]{biblatex}
    \DeclareLanguageMapping{american}{american-apa}
    \addbibresource{vcr.bib}
\usepackage{datetime}
\usepackage{url}

@@

\SetWatermarkAngle{0} % Angle at which the watermark text is drawn
\SetWatermarkColor{red} % Color of the watermark
%\SetWatermarkLightness{?real?} Lightness of the watermark text (1=white, 0=black)
%\SetWatermarkFontSize{?length?} Font size of the watermark text
\SetWatermarkScale{0.5} % Scaling of the watermark text 
%\SetWatermarkHorCenter{?length?} Horizontal center of watermark text 
\SetWatermarkVerCenter{2cm} % Vertical center of watermark text
\SetWatermarkText{Internal Draft} %Watermark text

\title{Vagueness as Cost Reduction}
\shorttitle{Vagueness as Cost Reduction}
\twoauthors{Matt Green}{Kees van Deemter}
\leftheader{Green, van Deemter} % for even-page headers in jou mode
\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
\abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations. We report a series of experiments that aims to separate the utility of vagueness from the utility of other factors that come together with vagueness. We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to vagueness itself.}
\keywords{keyword 1, keyword2, keyword3, keyword4}
\authornote{Contact author is Matt Green mjglab@gmail.com} % The Author Note, containing contact information, acknowledgements, etc.
\note{\date{\today~\currenttime}} % Notation of manuscript date or other information de- sired beneath affiliation line
\journal{Aimed at Language, Cognition, and Neuroscience} % typeset in the top left header of page 1 (jou and doc modes only)
\volume{volume info} % Volume, number, pages; typeset in the top left header in jou and doc modes, underneath the content of \journal
\ccoppy{Copyright notice etc} % Copyright notice, etc.; typeset in the top right header of page 1 (jou and doc modes only)
\copnum{Additional text} % Any additional text needed;  typeset in the top right header in jou and doc modes, underneath the content of \ccoppy

\begin{document}
\maketitle %
%\section{Introduction} % APA does not use Introduction header
Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons (see e.g. the entry ``vague'' in \cite{penguin}.) 
In most academic use though, the word `vagueness' has a more particular meaning. Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \textcite{EgreKlinedinst}, among many others).  The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}. A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not. The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.

@@

One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values. However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores. The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores. This ``illusion of proximity'' \cite[][p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance. In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}. These results are interesting, and of obvious potential practical importance. We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1996vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases. In the terminology of \textcite{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}. Furthermore, the experiments of \textcite{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.

>>>>

\documentclass[
a4paper 
%, man
, doc
%, jou
%, draftfirst 
, longtable
]{apa6} 
\usepackage{rotating} 
                          % use this in jou and doc modes only, with \sideways. 
                          % In man mode use the documentclass option longtable and do not load rotating.
\usepackage{csquotes}  
\usepackage[american]{babel}                     
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, doi=false, isbn=false]{biblatex}
    \DeclareLanguageMapping{american}{american-apa}
    \addbibresource{vcr.bib}
\usepackage{datetime}
\usepackage{url}

@@

%\SetWatermarkAngle{0} % Angle at which the watermark text is drawn
%\SetWatermarkColor{red} % Color of the watermark
%\SetWatermarkLightness{0.7} %Lightness of the watermark text (1=white, 0=black)
%%\SetWatermarkFontSize{?length?} Font size of the watermark text
%\SetWatermarkScale{0.5} % Scaling of the watermark text 
%%\SetWatermarkHorCenter{?length?} Horizontal center of watermark text 
%\SetWatermarkVerCenter{2cm} % Vertical center of watermark text
%\SetWatermarkText{Internal Draft} %Watermark text

\title{Vagueness as Cost Reduction}
\shorttitle{Vagueness as Cost Reduction}
\twoauthors{Matt Green}{Kees van Deemter}
\leftheader{Green, van Deemter} % for even-page headers in jou mode
\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
\abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations. We report a series of experiments that aims to separate the utility of vagueness from the utility of other factors that come together with vagueness. We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to vagueness itself.}
\keywords{keyword 1, keyword2, keyword3, keyword4}
\authornote{Contact author is Matt Green: \textsf{mjglab@gmail.com}} 
% The Author Note, containing contact information, acknowledgements, etc.
\note{This version compiled at: \textsf{\currenttime~\today}} 
% Notation of manuscript date or other information desired beneath affiliation line
\journal{Aimed at:}
% typeset in the top left header of page 1 (jou and doc modes only)
\volume{\textsc{Language, Cognition, and Neuroscience}} 
% Volume, number, pages; typeset in the top left header in jou and doc modes, underneath the content of \journal
\ccoppy{Draft date:}
% Copyright notice, etc.; typeset in the top right header of page 1 (jou and doc modes only)
\copnum{\today}
% Any additional text needed;  typeset in the top right header in jou and doc modes, underneath the content of \ccoppy
\begin{document}
\longdate
\maketitle % no blank line
%\section{Introduction} % APA does not use Introduction header
Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons (see e.g. the entry ``vague'' in \cite{penguin}.) 
In most academic use though, the word `vagueness' has a more particular meaning. Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness} (a similar definition can be found in \textcite{EgreKlinedinst}, among many others).  The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}. A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not. The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.

@@

One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values. However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores. The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores. This ``illusion of proximity'' \cite[][p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance. In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}. These results are interesting, and of obvious potential practical importance. We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases. In the terminology of \textcite{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}. Furthermore, the experiments of \textcite{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
